### 2013-09-22 structured_prediction/paper.tex
phi/psi
fix psi/Psi/phi/Phi mess :D

For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
\begin{equation}
    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
\end{equation}
The parameters $\theta$ are learned by minimizing the loss-based soft-margin
objective
\begin{equation}\eqlabel{learning_equation}
    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
\end{equation}
with regularization parameter $C$.\pagebreak\\
Here, $\ell$ is a hinge-loss like upper bound
on the empirical $\Delta$-risk:
\begin{equation}\eqlabel{loss_augmentation}
    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+.
\end{equation}
This is an instance of regularized empirical risk minimization, with a
piecewise linear, convex upper bound on the loss. 
Finding the $y$ that corresponds to a maximum in \Eqref{loss_augmentation} is
a central part of all maximum-margin based learning algorithms, and is referred to
as loss-augmented prediction. For complex models, such as the ones used for
image segmentation, this optimization often dominates the learning process in
terms of computational complexity.
Therefore, it is often desirable to find learning algorithms that converge with
as little optimizations of the loss-augmented prediction problem as possible.

@@

Arguably the most straight-forward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can becomputed simply by solving the loss-augmented prediction problem:
\begin{align*}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align*}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
In practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopt the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}.
\end{equation}

@@

\begin{algorithm*}[t]
    \caption{$n$-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for } i=1,\dots,k$
        \Repeat
            \For {i=1, \dots, k}
                \State
                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
                \State \hspace{-3mm}$\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
                    \State
                    \vspace{-15mm}
                    \begin{align*}
                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle \argmin_{\theta, \xi_1, \dots, \xi_k}& \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
                        \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:\\
                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
                    \end{align*}
                \EndIf
            \EndFor
            \vspace{-10mm}
            \Until no $\W_i$ changes anymore.
        \end{algorithmic}
    \end{doublespacing}
    \end{algorithm*}

@@

\begin{algorithm*}[t]
    \caption{$1$-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $\xi$
        \State $\W \leftarrow \emptyset$
        \Repeat
            \State 
            \vspace{-5mm}
            \begin{align*}
            (\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi}&\frac{||\theta||}{2}^2 + C \xi\\
            \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
             &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
            \end{align*}
            \For {i=1, \dots, k}
                \State
                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
            \EndFor
            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
    \end{algorithmic}
\end{algorithm*}

@@

\subsection{The SZLJSP Algorithm}\seclabel{dual_coordinate_descent}
\begin{algorithm*}[t]
    \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$
        \State  $\displaystyle \theta_0^{\vphantom{i}}, \theta_0\hoch{i}, \bar{\theta}_0^{\vphantom{i}} \leftarrow \mathbf{0},\quad \ell_0^{\vphantom{i}}, \ell_0\hoch{i}, t \leftarrow 0 $
        \Repeat
            \State $t \leftarrow t + 1$
            \State Pick $i$ uniformly at random from $\{1, \dotsc, k\}$.
            \State Perform loss-augmented prediction on sample $i$:
            \Statex[2]   $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \Delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$
            \State Compute parameter and loss updates based on sample $i$:
            \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \Phi(x, \hat{y})$
            \Statex[2]     $\ell_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
            \State Compute optimum step size $\eta$:
            \Statex[2]    $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
            \State Update per-sample parameters and loss estimate:
            \Statex[2]    $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$
            \Statex[2]    $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
            \State Update global parameters and loss estimate:
            \Statex[2]   $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$
            \Statex[2]   $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
            \State Compute the weighted running average:
            \Statex[2] $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
        \Until $(\theta - \theta_s)^T\theta - \ell + \ell_s \leq \epsilon$
        \Statex where $\theta_s$ and $\ell_s$ are recomputed over the whole dataset.
    \end{algorithmic}
    \end{doublespacing}
\end{algorithm*}

>>>>

For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
\begin{equation}
    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
\end{equation}
The parameters $\theta$ are learned by minimizing the loss-based soft-margin
objective
\begin{equation}\eqlabel{learning_equation}
    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
\end{equation}
with regularization parameter $C$.\pagebreak\\
Here, $\ell$ is a hinge-loss like upper bound
on the empirical $\Delta$-risk:
\begin{equation}\eqlabel{loss_augmentation}
    \ell(x\hoch{i}, y\hoch{i}, \theta) = [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y) - \theta^T \Phi(x\hoch{i}, y\hoch{i}]_+.
\end{equation}
This is an instance of regularized empirical risk minimization, with a
piecewise linear, convex upper bound on the loss. 
Finding the $y$ that corresponds to a maximum in \Eqref{loss_augmentation} is
a central part of all maximum-margin based learning algorithms, and is referred to
as loss-augmented prediction. For complex models, such as the ones used for
image segmentation, this optimization often dominates the learning process in
terms of computational complexity.
Therefore, it is often desirable to find learning algorithms that converge with
as little optimizations of the loss-augmented prediction problem as possible.

@@

Arguably the most straight-forward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
Given a model through $\Phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can becomputed simply by solving the loss-augmented prediction problem:
\begin{align*}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y)
\end{align*}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
In practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopt the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}.
\end{equation}

@@

\begin{algorithm*}[t]
    \caption{$n$-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for } i=1,\dots,k$
        \Repeat
            \For {i=1, \dots, k}
                \State
                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})] $
                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
                \State \hspace{-3mm}$\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
                    \State
                    \vspace{-15mm}
                    \begin{align*}
                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle \argmin_{\theta, \xi_1, \dots, \xi_k}& \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
                        \text{s.t. }& \text{for } i=1,\dots,k\ \forall \hat{y} \in \W_i:\\
                    &\theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
                    \end{align*}
                \EndIf
            \EndFor
            \vspace{-10mm}
            \Until no $\W_i$ changes anymore.
        \end{algorithmic}
    \end{doublespacing}
    \end{algorithm*}

@@

\begin{algorithm*}[t]
    \caption{$1$-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $\xi$
        \State $\W \leftarrow \emptyset$
        \Repeat
            \State 
            \vspace{-5mm}
            \begin{align*}
            (\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi}&\frac{||\theta||}{2}^2 + C \xi\\
            \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
             &\theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
            \end{align*}
            \For {i=1, \dots, k}
                \State
                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
            \EndFor
            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y}\hoch{i})] $
        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
    \end{algorithmic}
\end{algorithm*}

@@

\subsection{The SZLJSP Algorithm}\seclabel{dual_coordinate_descent}
\begin{algorithm*}[t]
    \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$
        \State  $\displaystyle \theta_0^{\vphantom{i}}, \theta_0\hoch{i}, \bar{\theta}_0^{\vphantom{i}} \leftarrow \mathbf{0},\quad \ell_0^{\vphantom{i}}, \ell_0\hoch{i}, t \leftarrow 0 $
        \Repeat
            \State $t \leftarrow t + 1$
            \State Pick $i$ uniformly at random from $\{1, \dotsc, k\}$.
            \State Perform loss-augmented prediction on sample $i$:
            \Statex[2]   $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \Delta(y\hoch{i}, \hat{y}) - \theta^T [\Phi(x\hoch{i}, y\hoch{i}) - \Phi(x\hoch{i}, \hat{y})]$
            \State Compute parameter and loss updates based on sample $i$:
            \Statex[2]     $\theta_s \leftarrow \frac{C}{n} \Phi(x, \hat{y})$
            \Statex[2]     $\ell_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
            \State Compute optimum step size $\eta$:
            \Statex[2]    $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
            \State Update per-sample parameters and loss estimate:
            \Statex[2]    $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$
            \Statex[2]    $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
            \State Update global parameters and loss estimate:
            \Statex[2]   $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$
            \Statex[2]   $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
            \State Compute the weighted running average:
            \Statex[2] $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
        \Until $(\theta - \theta_s)^T\theta - \ell + \ell_s \leq \epsilon$
        \Statex where $\theta_s$ and $\ell_s$ are recomputed over the whole dataset.
    \end{algorithmic}
    \end{doublespacing}
\end{algorithm*}
